An Analysis of the Feasibility and Performance of BitTorrent as a Data Transfer Protocol in Peer to Peer Backup Systems

### 2014-04-15 paper/paper.tex
Paper: fixes...

The data layer is responsible for copying the data to be backed up from the peer to the node. This includes the quick communication by the peer to the node informing it that it is a replicant node for a given file.

Our main contribution in this paper, and the only requirement of the data layer, is that it uses BitTorrent as the data transport portocol. By doing so, we increase the speed of data recovery because the peer can download the data from all replicant nodes simultaneously.

@@

The data that is transported via the data layer is encrypted so that the replicant nodes cannot read it. The strength of the encryption is left unspecified as a means of future-proofing this specification.

@@

%If less than X number of replicas, a peer should make a new replica on a node that has yet to hold a replica.
For users of our system to take advantage of the speed increases offered by the BitTorrent protocol, a sufficient number of replicas need to be maintained in the overlay network (for a comprehensive explanation of how the BitTorrent protocol works, see section \ref{sec:TheBitTorrentProtocol}). Since the P2P network will be composed of ordinary network users, backups that are created may not always be available. Therefore, it is not enough to rely on the backups that are created when a file is first backed up; in the worst-case scenario, all of the nodes that a user has created replicas on have permanently left the network, leaving the user with no way to retrieve any backup. Even if one or two replica nodes are online when the backups are needed,  the user will not see much of a speed benefit from the BitTorrent protocol, as opposed to all replicas being online.

% TODO: please pay extra attention during the review of this paragraph; I had to work in references to the challenge system after writing it, so some things might sound awkward.
To solve this problem, peers in our system periodically check on the nodes that they are using as backup locations in order to gauge how available they are. During normal program operation, each peer tests the availability of its backup locations (along with a challenge mechanism, which is described in section \ref{subsec:ChallengeMechanism_Fairness}) to see whether or not they are online. If a backup node is found to be online and it passes the challenge, then the peer does not need to do anything; it will continue to think that it is reliable. If the node is offline or fails the challenge, however, the peer will take note that it was unavailable during an availability test, and that the node might be unreliable in general. It would be unfair to immediately conclude that the node is unreliable, since it may have been temporarily offline (for example, a user temporarily shutting down his/her computer). Therefore, a node needs to fail the ping test several times before the peer that uses it concludes that it is unreliable. It is possible that a node gets unlucky and happens to be offline during all checks, even though that it is online during a different part of the day. However, if this is the case, then the node is not frequently available for that node, and should be seen as unreliable.

% TODO: rework this paragraph to make it flow better now that the two paragraphs have been combined
When a peer determines that a node is unreliable, it does two things. The first is that it adds its own ID to the node's blacklist in the DHT. The second is to move its replica to a new node. The peer sends an update to the DHT that it is removing the replica on the unreliable node. This accomplishes two things: all other peers in the network can see that the node is storing less data, and the node can see that it no longer needs to store the replica, so it can remove it the next time it checks the DHT for any changes. The peer will then look for a new node to put a replica on, using the same process that was used for creating the original backups (see section \ref{sec:TheMetadataLayer}). Since this process uses a randomly generated salt to help pick the ID of the new node, each node has an equal chance of being selected, assuming that it has enough space to store the replica and is not blacklisted by too many peers. This helps to prevent some nodes from being left out of the selection process.

\section{Fairness} \label{sec:Fairness}
% People store proportional amount of data compared to how much they are backing up
To ensure that each node is being fair (see section \ref{sec:BackgroundFairness}), each node must contribute storage space proportional to the amount that they are backing up. Since a peer will have multiple replicas of each file, the amount of space that a peer offers must be much larger than the amount of unique data that it is backing up. If $n$ backups are maintained for each file and $F$ represents the set of files that a peer backs up, the total amount of space $t$ that the peer must offer can be expressed as:
\begin{equation}
t=n\sum\limits_{i=1}^{|F|} size(F_i)
\end{equation}
There still exists one problem with this mechanism: when the network first starts up, no peer has created any backups, so no one is obligated to store data for other peers. To solve this, our system requires that each peer is obligated to take at least $s$ bytes of data from other peers, regardless of how much data it has backed up. To include this, $t$ can be reexpressed as:
\begin{equation}
t=max(s,n\sum\limits_{i=1}^{|F|} size(F_i))
\end{equation}
When a peer finds a node that has a value of $t$ greater than the size of the file it is trying to replicate, it can choose that node as a replication site, knowing that it will be able to accept the data.

>>>>

The data layer is responsible for copying the data to be backed up from the peer to the node. This includes the quick communication by the peer to the node informing it that it is a replication node for a given file.

Our main contribution in this paper, and the only requirement of the data layer, is that it uses BitTorrent as the data transport protocol. By doing so, we increase the speed of data recovery because the peer can download the data from all replication nodes simultaneously.

@@

The data that is transported via the data layer is encrypted so that the replication nodes cannot read it. The strength of the encryption is left unspecified as a means of future-proofing this specification.

@@

%If less than X number of replicas, a peer should make a new replica on a node that has yet to hold a replica. (is this a TODO?)
For users of our system to take advantage of the speed increases offered by the BitTorrent protocol, a sufficient number of replicas need to be maintained in the overlay network (for a comprehensive explanation of how the BitTorrent protocol works, see section \ref{sec:TheBitTorrentProtocol}). Since the P2P network will be composed of ordinary network users, backups that are created may not always be available. Therefore, it is not enough to rely on the backups that are created when a file is first backed up; in the worst-case scenario, all of the nodes on which a user has created replicas have permanently left the network, leaving the user with no way to retrieve any backup. If one or two replication nodes are online when the backups are needed, the user will be able to recover their file but will not see much of a speed benefit from the BitTorrent protocol, as opposed to when all replication nodes are online.

To solve this problem, peers in our system periodically check on the nodes that they are using as backup locations in order to gauge how available they are. During normal program operation, each peer tests the availability of its replication nodes via a challenge mechanism, described in section \ref{subsec:ChallengeMechanism_sec:Fairness}. The challenge mechanism not only ensures the node is online, but also confirms it is storing the peer's data. If the replication node is online and passes the challenge, the peer does not need to do anything; it will continue to (accurately) think that the node is reliable. If the node fails the challenge (by being offline or not storing the user's data), however, the peer will take note that the node was unreliable in this instance. It would be unfair to immediately conclude that the node is consistently unreliable, since it may have been temporarily offline (for example, a user temporarily shutting down his/her computer). Therefore, a node needs to fail the challenge mechanism several times before the peer concludes that it is consistently unreliable. It is possible that a node gets unlucky and happens to be offline during all checks, even though it is online during a different part of the day. If this is the case, then the node is not frequently available for that peer, and should be seen as consistently unreliable from the peer's perspective.

When a peer determines that a node is consistently unreliable, it does two things. The first is that it adds its own ID to the node's blacklist in the DHT. The second is to move its replica to a new node. The peer sends an update to the DHT that it is removing the replica on the unreliable node. This accomplishes two things: all other peers in the network can see that the node is storing less data, and the node can see that it no longer needs to store the replica, so it can remove it the next time it checks the DHT for any changes. The peer will then look for a new node to put a replica on, using the same process that was used for creating the original backups (see section \ref{sec:TheMetadataLayer}). Since this process uses a randomly generated salt to help pick the ID of the new node, each node has an equal chance of being selected, assuming that it has enough space to store the replica and is not blacklisted by too many peers. This helps to prevent some nodes from being left out of the selection process. % TODO this last sentence can be removed without loss of information, but it might mean that the salt explanation in the crypto section was for nothing.

\section{Fairness} \label{sec:Fairness}
% People store proportional amount of data compared to how much they are backing up
% TODO we say "if n backups", but also refer to each backup as a "replica". we should be consistent. where else do we do this? ctrl-f replica
To ensure fairness (see section \ref{sec:BackgroundFairness} for background information), each node must contribute storage space proportional to the amount that they are backing up. Since a peer will have replicas of its file stored on multiple nodes, the amount of space that a peer offers\footnote{That is, offers to store when it assumes the role of a node} must be much larger than the amount of unique data that it is backing up. If $n$ backups are maintained for each file and $F$ represents the set of files that a peer backs up, the total amount of space $t$ that the peer must offer can be expressed as:
\begin{equation}
t=n\sum\limits_{i=1}^{|F|} size(F_i)
\end{equation}
There exists one problem with this mechanism: when the network first starts up, no peer has created any backups, so no one is obligated to store data for other peers. To solve this, our system requires that each peer is obligated to take at least $s$ bytes of data from other peers, regardless of how much data it has backed up. To include this, $t$ can be reexpressed as:
\begin{equation}
t=max(s,n\sum\limits_{i=1}^{|F|} size(F_i))
\end{equation}
When a peer finds a node that has a value of $t$ greater than the size of the file it is trying to replicate, it can choose that node as a replication site, knowing that it will be able to accept the data.
